---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# baymedr: BAYesian inference for MEDical designs in R

<!-- badges: start -->
<!-- badges: end -->

`baymedr` is an R package with the goal of providing researchers with functions for the computation of Bayes factors for common biomedical research designs. Implemented are functions to test the equivalence (`equiv_bf`), non-inferiority (`infer_bf`), and superiority (`super_bf`) of an intervention group (e.g., a new medication) compared to some form of control group (e.g., a placebo or an already existing medication). Bayes factors for these three tests can be computed based on raw data (arguments: x, y) or summary statistics (arguments: n_x, n_y, mean_x, mean_y, sd_x, sd_y), making it possible to reanalyze findings (e.g., from publications) without the need to obtain the raw data.

## Installation

You can install the latest development version of `baymedr` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("maxlinde/baymedr")
```

## Random example data

In order to demonstrate the three functions from `baymedr`, we create an example dataset (data). We need a control group "con" and an experimental group "exp" (condition). Further, random numbers, sampled from the normal distribution, within each group are created, serving as the dependent variable of interest (dv):

``` {r}
set.seed(1)

data <- data.frame(
  condition = rep(x = c("con", "exp"),
                  c(150, 180)),
  dv = c(rnorm(n = 150,
               mean = 7.3,
               sd = 3.4),
         rnorm(n = 180,
               mean = 8.9,
               sd = 3.1))
)

head(x = data,
     n = 10)
```

## The superiority test (`super_bf`)

With `super_bf` we can test the null hypothesis that the experimental group (e.g., a new medication) is not better than the control group (e.g., a placebo or existing medication). Because research practices diverge on whether to conduct a superiority test as a one-sided or two-sided test, two Bayes factors are obtained from `super_bf`, one for each research strategy.

We can use the raw data to compute Bayes factors:

``` {r}
library(baymedr)

super_bf(x = data$dv[data$condition == "con"],
         y = data$dv[data$condition == "exp"])
```

Alternatively, if the raw data are not available, we can use summary statistics to compute Bayes factors:

``` {r}
library(baymedr)

super_bf(n_x = 138,
         n_y = 179,
         mean_x = 18.3,
         mean_y = 21.9,
         sd_x = 3.3,
         sd_y = 4.0)
```

## The equivalence test (`equiv_bf`)

With `equiv_bf` we can test the null hypothesis that the control group (e.g., a placebo or existing medication) and the experimental group (e.g., a new medication) are equivalent. The alternative hypothesis is that the two groups are not equivalent.

We can use the raw data to compute the Bayes factor:

``` {r}
library(baymedr)

equiv_bf(x = data$dv[data$condition == "con"],
         y = data$dv[data$condition == "exp"])
```

Alternatively, if the raw data are not available, we can use summary statistics to compute the Bayes factor:

``` {r}
library(baymedr)

equiv_bf(n_x = 138,
         n_y = 179,
         mean_x = 18.3,
         mean_y = 21.9,
         sd_x = 3.3,
         sd_y = 4.0)
```

## The non-inferiority test (`infer_bf`)

With `infer_bf` we can test the null hypothesis that the experimental group (e.g., a new medication) is not better than the control group (e.g., a placebo or existing medication) minus a constant value (c). The alternative hypothesis is that the experimental group is better than the control group minus a constant value (c). Put more formally, the null hypothesis states that the true population effect size < -c, resulting in the point null hypothesis that the true population effect size = -c against the one-sided alternative hypothesis that the true population effect size > -c.

We can use the raw data to compute the Bayes factor:

``` {r}
library(baymedr)

infer_bf(x = data$dv[data$condition == "con"],
         y = data$dv[data$condition == "exp"],
         ni_margin = 2)
```

Alternatively, if the raw data are not available, we can use summary statistics to compute the Bayes factor:

``` {r}
library(baymedr)

infer_bf(n_x = 138,
         n_y = 179,
         mean_x = 18.3,
         mean_y = 21.9,
         sd_x = 3.3,
         sd_y = 4.0,
         ni_margin = 2)
```
