---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# baymedr: BAYesian inference for MEDical designs in R

`baymedr` is an R package with the goal of providing researchers with easy-to-use tools for the computation of Bayes factors for common biomedical research designs. Implemented are functions to test the equivalence (`equiv_bf()`), non-inferiority (`infer_bf()`), and superiority (`super_bf()`) of an experimental group (e.g., a new medication) compared to a control group (e.g., a placebo or an already existing medication) on a continuous dependent variable, as well as functions for simulating survival data (`coxph_data_sim()`) and calculating a Bayes factor for Cox proportional hazards models (`coxph_bf()`). A special focus of `baymedr` lies on a user-friendly interface, so that a wide variety or researchers (i.e., not only statisticians) can utilize `baymedr` for their analyses.

## Installation and attaching

To install `baymedr` use:

``` {r, eval = FALSE}
install.packages("baymedr")
```

You can install the latest development version of `baymedr` from [GitHub](https://github.com/), using the `devtools` package, with:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("maxlinde/baymedr")
```

Subsequently, you can load `baymedr`, so that it is ready to use:

``` {r, eval = FALSE}
library(baymedr)
```
